---
title: "SPP_Figures"
author: "Jesse Wood"
date: "July 17, 2018 -- update now with combo years!!"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#using model avg package

```{r}
library(AICcmodavg)
library(unmarked)
library(sjPlot)
library(sjmisc)
library(ggplot2)
setwd("C:/Users/woodj/Documents/GRAD SCHOOL - CLEMSON/Project-Specific/R work/USDA-songbirds/USDA-songbirds")
```

#predict with unmarked    # code from ESA 2014
  #remember, treat differently for those that are factors vs. those that are #s

#general structure:
fitList of top models (d<2) for SP
write data frame including all important variables -
       # concat factor (levels) if necessary
      # set interest variable to max/min
      # but set every other variable besides one of interest to 0
define any factors with categorical names
predict function on the new matrix, type = state
plot matrix & 95% CI lines
  #barplot with error bars code if need be - see bottom

```{r}
#PIWA
piwa.abund<- csvToUMF("piwa_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(piwa.abund)= scale (obsCovs(piwa.abund))
sc <- siteCovs(piwa.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(piwa.abund) <- sc

landscape5.piwa <- pcount(~ Jdate + Time 
                       ~ Evergreen5km + HighDev5km
                          , piwa.abund, mixture="P", K=40)

#evergreen5km from landscape 5km model
NewData.Evergreen5km  <- data.frame(Evergreen5km= seq(min(sc$Evergreen5km),max(sc$Evergreen5km), length=100), HighDev5km = 0)                             
piwa.est.evergreen5km <- predict(landscape5.piwa, type="state", newdata=NewData.Evergreen5km,appendData=TRUE)

plot(Predicted~ Evergreen5km, data=piwa.est.evergreen5km, ylim=c(0,16), type="l", lwd=3, xlab="Evergreen habitat within 5km of patch", ylab="Est. PIWA Abundance")
##95% confidence intervals
lines(lower~ Evergreen5km, data=piwa.est.evergreen5km,  type="l", lwd=3, col="darkgray")
lines(upper~ Evergreen5km, data=piwa.est.evergreen5km, type="l", lwd=3, col="darkgray")
```

```{r}
#PIWA
piwa.abund<- csvToUMF("piwa_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(piwa.abund)= scale (obsCovs(piwa.abund))
sc <- siteCovs(piwa.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(piwa.abund) <- sc

landscape30.piwa <- pcount(~ Jdate + Time 
                       ~ Evergreen30km + HighDev30km + Protected30km
                           , piwa.abund, mixture="P", K=40)

#High-Dev 30km
NewData.HighDev30km <- data.frame(Evergreen30km=0,
                      HighDev30km=seq(min(sc$HighDev30km),max(sc$HighDev30km), length=100),
                      Protected30km=0)
                                       
piwa.est.highdev30km <- predict(landscape30.piwa, type="state", newdata=NewData.HighDev30km,appendData=TRUE)

plot(Predicted~ HighDev30km, data=piwa.est.highdev30km, ylim=c(0,15), type="l", lwd=3, xlab="Highly Developed Area within 30km of patch", ylab="Est. PIWA Abundance")
##95% confidence intervals
lines(lower~ HighDev30km, data=piwa.est.highdev30km,  type="l", lwd=3, col="darkgray")
lines(upper~ HighDev30km, data=piwa.est.highdev30km, type="l", lwd=3, col="darkgray")
```

```{r}
#PRWA - landscape 5km model, high development variable
prwa.abund<- csvToUMF("prwa_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(prwa.abund)= scale (obsCovs(prwa.abund))
sc <- siteCovs(prwa.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(prwa.abund) <- sc
landscape5.prwa <- pcount(~ Jdate
                          ~ Evergreen5km + Grass5km + HighDev5km + Schrubs5km
                          , prwa.abund, mixture="NB", K=60)

NewData.HighDev5km  <- data.frame(Evergreen5km=0, Grass5km=0, HighDev5km=seq(min(sc$HighDev5km),max(sc$HighDev5km),length=100), Schrubs5km=0)
prwa.est.highdev5 <- predict(landscape5.prwa, type="state",newdata=NewData.HighDev5km,appendData=TRUE)

plot(Predicted~ HighDev5km, data=prwa.est.highdev5, ylim=c(0,5), type="l", lwd=3,
     xlab="Highly developed area within 5km of patch", ylab="Est. PRWA Abundance")
##95% confidence intervals
lines(lower~ HighDev5km, data=prwa.est.highdev5,  type="l", lwd=3, col="darkgray")
lines(upper~ HighDev5km, data=prwa.est.highdev5, type="l", lwd=3, col="darkgray")
```

# YBCH - landscape 1km model best, 5km model second best - but no sigs!

```{r}
#INBU
inbu.abund<- csvToUMF("inbu_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(inbu.abund)= scale (obsCovs(inbu.abund))
sc <- siteCovs(inbu.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(inbu.abund) <- sc
disturbance.inbu <- pcount(~ Jdate
                           ~ TimeSinceB + TimeSinceT
                           , inbu.abund, mixture="P", K=40)

#INBU - disturbance model, Time Since Burn variable

#experimentation#
newData.inbu<-data.frame(TimeSinceT=0, TimeSinceB=0:20)
newdata.inbu.est.timeburn<-predict(disturbance.inbu, type="state", newdata=newData.inbu, appendData=TRUE, 2)

plot(Predicted~ TimeSinceB, data=newdata.inbu.est.timeburn, ylim=c(0,6), type="l", lwd=3,
     xlab="Time Since Burn (years)", ylab="Est. INBU Abundance")
##95% confidence intervals
lines(lower~ TimeSinceB, data=newdata.inbu.est.timeburn,  type="l", lwd=3, col="darkgray")
lines(upper~ TimeSinceB, data=newdata.inbu.est.timeburn, type="l", lwd=3, col="darkgray")

## keep below
NewData.TsinceB  <- data.frame(TimeSinceT= 0, TimeSinceB=seq(min(sc$TimeSinceB),max(sc$TimeSinceB), length=100))
inbu.est.timeburn <- predict(disturbance.inbu, type="state",
                                newdata=NewData.TsinceB,appendData=TRUE)

plot(Predicted~ TimeSinceB, data=inbu.est.timeburn, ylim=c(0,5), type="l", lwd=3,
     xlab="Time Since Burn", ylab="Est. INBU Abundance")
##95% confidence intervals
lines(lower~ TimeSinceB, data=inbu.est.timeburn,  type="l", lwd=3, col="darkgray")
lines(upper~ TimeSinceB, data=inbu.est.timeburn, type="l", lwd=3, col="darkgray")
```
#INBU - nothing significant for life history model

```{r}
#INBU - treatment model, 1B & 3B significant variable

treatment.inbu <- pcount(~ Jdate
                         ~ Treatment + Nthins
                         , inbu.abund, mixture ="P", K=40)

##
error.bar <- function(x, y, upper, lower=upper, length=0.1,...){
  if(length(x) != length(y) | length(y) !=length(lower) | length(lower) != length(upper))
    stop("vectors must be same length")
  arrows(x,y+upper, x, y-lower, angle=90, code=3, length=length, ...)
}
##

NewData.matrix  <- data.frame(Treatment=factor(c("0B","1B","2B","3B")),
                              Nthins=0)
inbu.est.matrix <- predict(treatment.inbu, type="state",
                           newdata=NewData.matrix,appendData=TRUE)

data.mean<-inbu.est.matrix$Predicted
data.sd<-inbu.est.matrix$SE

bar.p <-barplot(data.mean,
                names.arg=c("0B","1B","2B","3B"),
                ylim = c(0, 8), ylab="Est. INBU Abundance", xlab="Treatment",
                #cex.names = 1.5, cex.axis=1.5, cex.lab=1.5, 
                col="darkblue")

error.bar(bar.p,data.mean,data.sd) #sd

```
```{r}
#BHNU
bhnu.abund<- csvToUMF("bhnu_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(bhnu.abund)= scale (obsCovs(bhnu.abund))
sc <- siteCovs(bhnu.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(bhnu.abund) <- sc
lh.bhnu <- pcount(~ Time
                  ~ Ccover + Age + Nsnags + TreeHt + NP_over_20cm + Rel_HW2P_canopy
                  , bhnu.abund, mixture="P", K=40)

#BHNU - life history model, relative hardwood to pine canopy cover variable

NewData.RelHW2Pcanopy  <- data.frame(Ccover=0,Age=0,Nsnags=0,TreeHt=0,NP_over_20cm=0,
                              Rel_HW2P_canopy=seq(min(sc$Rel_HW2P_canopy),max(sc$Rel_HW2P_canopy), length=100))
bhnu.est.relhw2p <- predict(lh.bhnu, type="state",
                                newdata=NewData.RelHW2Pcanopy,appendData=TRUE)

plot(Predicted~ Rel_HW2P_canopy, data=bhnu.est.relhw2p, ylim=c(0,50), type="l", lwd=3,
     xlab="Increasing Hardwood to Pine Ratio in Canopy", ylab="Est. BHNU Abundance")
##95% confidence intervals
lines(lower~ Rel_HW2P_canopy, data=bhnu.est.relhw2p,  type="l", lwd=3, col="darkgray")
lines(upper~ Rel_HW2P_canopy, data=bhnu.est.relhw2p, type="l", lwd=3, col="darkgray")

```

```{r}
#BHNU - life history model, large pines variable

NewData.NPover20cm  <-data.frame(Ccover=0,Age=0,Nsnags=0,TreeHt=0, Rel_HW2P_canopy=0, NP_over_20cm=seq(min(sc$NP_over_20cm),max(sc$NP_over_20cm), length=100))
bhnu.est.NPover20cm <- predict(lh.bhnu, type="state",
                                newdata=NewData.NPover20cm,appendData=TRUE)

plot(Predicted~ NP_over_20cm, data=bhnu.est.NPover20cm, ylim=c(0,50), type="l", lwd=3,
     xlab="Number of large pines (>20cm diameter)", ylab="Est. BHNU Abundance")
##95% confidence intervals
lines(lower~ NP_over_20cm, data=bhnu.est.NPover20cm,  type="l", lwd=3, col="darkgray")
lines(upper~ NP_over_20cm, data=bhnu.est.NPover20cm, type="l", lwd=3, col="darkgray")

```

```{r}
# EABL
eabl.abund<- csvToUMF("eabl_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(eabl.abund)= scale (obsCovs(eabl.abund))
sc <- siteCovs(eabl.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(eabl.abund) <- sc
landmetrics.eabl <- pcount (~ Jdate
                          ~ Parea + ShapeIndex
                        , eabl.abund, mixture="P",K=80)

#EABL - land metrics model, shape index variable

NewData.shapeindex  <-data.frame(Parea=0,ShapeIndex=seq(min(sc$ShapeIndex),max(sc$ShapeIndex), length=100))
eabl.est.shapeindex <- predict(landmetrics.eabl, type="state",
                                newdata=NewData.shapeindex,appendData=TRUE)

plot(Predicted~ ShapeIndex, data=eabl.est.shapeindex, ylim=c(0,6), type="l", lwd=3,
     xlab="Shape Index (increasingly complex, less square)", ylab="Est. EABL Abundance")
##95% confidence intervals
lines(lower~ ShapeIndex, data=eabl.est.shapeindex,  type="l", lwd=3, col="darkgray")
lines(upper~ ShapeIndex, data=eabl.est.shapeindex, type="l", lwd=3, col="darkgray")

```

#NOBO - a ton
```{r}
#NOBO - life history model, basal area variable

nobo.abund<- csvToUMF("nobo_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(nobo.abund)= scale (obsCovs(nobo.abund))
sc <- siteCovs(nobo.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(nobo.abund) <- sc
lh.nobo <- pcount(~ Jdate + Wind + Sky + Noise +Time
                  ~ BA + Ldepth + HW_dens_1050 + FG_shrub + Age
                  , nobo.abund, mixture="NB", K=80)
lh.nobo
confint(lh.nobo, type="state",method="normal")


NewData.BA  <-data.frame(BA=seq(min(sc$BA),max(sc$BA),length=100),
                         Age=0,FG_shrub=0,Ldepth=0,HW_dens_1050=0)
nobo.est.ba <- predict(lh.nobo, type="state",
                                newdata=NewData.BA,appendData=TRUE)

plot(Predicted~ BA, data=nobo.est.ba, ylim=c(0,10), type="l", lwd=3,
     xlab="Basal Area (10-factor)", ylab="Est. NOBO Abundance")
##95% confidence intervals
lines(lower~ BA, data=nobo.est.ba,  type="l", lwd=3, col="darkgray")
lines(upper~ BA, data=nobo.est.ba, type="l", lwd=3, col="darkgray")

```

```{r}
#NOBO - life history model, forbes&grasses_shrub ht variable

NewData.FGshrub  <-data.frame(BA=0,Age=0,FG_shrub=seq(min(sc$FG_shrub),max(sc$FG_shrub),length=100),
                         Ldepth=0,HW_dens_1050=0)
nobo.est.fgshrub <- predict(lh.nobo, type="state",
                                newdata=NewData.FGshrub,appendData=TRUE)

plot(Predicted~ FG_shrub, data=nobo.est.fgshrub, ylim=c(0,30), type="l", lwd=3,
     xlab="Herbaceous Vegetation (mid-level)", ylab="Est. NOBO Abundance")
##95% confidence intervals
lines(lower~ FG_shrub, data=nobo.est.fgshrub,  type="l", lwd=3, col="darkgray")
lines(upper~ FG_shrub, data=nobo.est.fgshrub, type="l", lwd=3, col="darkgray")

```

```{r}
#NOBO - local model, canopy cover variable

local.nobo <- pcount(~ Jdate + Wind + Sky + Noise +Time
                     ~ Ccover + TreeHt + Ldepth
                     , nobo.abund, mixture="NB", K=80)

NewData.ccover  <-data.frame(Ccover=seq(min(sc$Ccover),max(sc$Ccover),length=100),TreeHt=0,Ldepth=0)
nobo.est.ccover <- predict(local.nobo, type="state",
                                newdata=NewData.ccover,appendData=TRUE)

plot(Predicted~ Ccover, data=nobo.est.ccover, ylim=c(0,40), type="l", lwd=3,
     xlab="% Canopy Cover", ylab="Est. NOBO Abundance")
##95% confidence intervals
lines(lower~ Ccover, data=nobo.est.ccover,  type="l", lwd=3, col="darkgray")
lines(upper~ Ccover, data=nobo.est.ccover, type="l", lwd=3, col="darkgray")

```

```{r}
#NOBO - landscape 1km model, evergreen variable
landscape1.nobo <- pcount(~ Jdate + Wind + Sky + Noise +Time
                       ~  Grass1km + HighDev1km + Schrubs1km + Evergreen1km + Ag1km
                          , nobo.abund, mixture="NB", K=80)

NewData.Evergreen1km  <- data.frame(Evergreen1km=seq(-2,3,length=100), Grass1km=0,HighDev1km=0,Schrubs1km=0,Ag1km=0)
nobo.est.evergreen1 <- predict(landscape1.nobo, type="state",
                              newdata=NewData.Evergreen1km,appendData=TRUE)

plot(Predicted~ Evergreen1km, data=nobo.est.evergreen1, ylim=c(0,5), type="l", lwd=3,
     xlab="Evergreen habitat within 1km of patch", ylab="Est. NOBO Abundance")
##95% confidence intervals
lines(lower~ Evergreen1km, data=nobo.est.evergreen1,  type="l", lwd=3, col="darkgray")
lines(upper~ Evergreen1km, data=nobo.est.evergreen1, type="l", lwd=3, col="darkgray")
```

```{r}
#NOBO - treatment model, #Thins variable
treatment.nobo <- pcount(~ Jdate + Wind + Sky + Noise +Time
                         ~ Nburns + Nthins
                         , nobo.abund, mixture ="NB", K=80)  #this one I made Nburns instead of treatment for ones with many many burns

NewData.Nthins  <- data.frame(Nburns= 0, Nthins=seq(min(sc$Nthins),max(sc$Nthins), length=100))
nobo.est.nthins <- predict(treatment.nobo, type="state",
                                newdata=NewData.Nthins,appendData=TRUE)

plot(Predicted~ Nthins, data=nobo.est.nthins, ylim=c(0,10), type="l", lwd=3,
     xlab="Number of thinnings", ylab="Est. NOBO Abundance")
##95% confidence intervals
lines(lower~ Nthins, data=nobo.est.nthins,  type="l", lwd=3, col="darkgray")
lines(upper~ Nthins, data=nobo.est.nthins, type="l", lwd=3, col="darkgray")
```


```{r}
#NOBO - landscape 5km model, evergreen variable
landscape5.nobo <- pcount(~ Jdate + Wind + Sky + Noise +Time
                       ~ Grass5km + HighDev5km + Schrubs5km + Evergreen5km
                          , nobo.abund, mixture="NB", K=80)

NewData.Evergreen5km  <- data.frame(Evergreen5km=seq(min(sc$Evergreen5km),max(sc$Evergreen5km), length=100),Grass5km=0,HighDev5km=0,Schrubs5km=0)
nobo.est.evergreen5 <- predict(landscape5.nobo, type="state",
                              newdata=NewData.Evergreen5km,appendData=TRUE)

plot(Predicted~ Evergreen5km, data=nobo.est.evergreen5, ylim=c(0,30), type="l", lwd=3,
     xlab="Evergreen habitat within 5km of patch (acres)", ylab="Est. NOBO Abundance")
##95% confidence intervals
lines(lower~ Evergreen5km, data=nobo.est.evergreen5,  type="l", lwd=3, col="darkgray")
lines(upper~ Evergreen5km, data=nobo.est.evergreen5, type="l", lwd=3, col="darkgray")
```



#EAWP - a few
```{r}
#EAWP - landscape 500m model, evergreen variable
eawp.abund<- csvToUMF("eawp_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(eawp.abund)= scale (obsCovs(eawp.abund))
sc <- siteCovs(eawp.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)])
siteCovs(eawp.abund) <- sc
landscape500.eawp <- pcount(~ Wind + Sky
                        ~ Evergreen500m + Ag500m + HighDev500m
                            , eawp.abund, mixture="P", K=40)

NewData.Evergreen500m  <- data.frame(Evergreen500m=seq(min(sc$Evergreen500m),max(sc$Evergreen500m), length=100),Ag500m=0,HighDev500m=0)
eawp.est.evergreen500 <- predict(landscape500.eawp, type="state",
                              newdata=NewData.Evergreen500m,appendData=TRUE)

plot(Predicted~ Evergreen500m, data=eawp.est.evergreen500, ylim=c(0,5), type="l", lwd=3,
     xlab="Evergreen habitat within 500m of patch", ylab="Est. EAWP Abundance")
##95% confidence intervals
lines(lower~ Evergreen500m, data=eawp.est.evergreen500,  type="l", lwd=3, col="darkgray")
lines(upper~ Evergreen500m, data=eawp.est.evergreen500, type="l", lwd=3, col="darkgray")
```
```{r}
#EAWP - local model, tree height variable
local.eawp <- pcount(~ Wind + Sky
                     ~ Ccover + TreeHt + Ldepth
                     , eawp.abund, mixture="P", K=40)

NewData.TreeHt  <- data.frame(TreeHt=seq(min(sc$TreeHt),max(sc$TreeHt), length=100),
                                    Ccover=0,Ldepth=0)
eawp.est.treeht <- predict(local.eawp, type="state",
                              newdata=NewData.TreeHt,appendData=TRUE)

plot(Predicted~ TreeHt, data=eawp.est.treeht, ylim=c(0,5), type="l", lwd=3,
     xlab="Tree height", ylab="Est. EAWP Abundance")
##95% confidence intervals
lines(lower~ TreeHt, data=eawp.est.treeht,  type="l", lwd=3, col="darkgray")
lines(upper~ TreeHt, data=eawp.est.treeht, type="l", lwd=3, col="darkgray")
```

```{r}
#RBWO - landscape 1km model, ag variable
rbwo.abund<- csvToUMF("rbwo_abund.csv", long = FALSE, type = "unmarkedFramePCount")
obsCovs(rbwo.abund)= scale (obsCovs(rbwo.abund))
sc <- siteCovs(rbwo.abund)
sc[,c(5:74)] <- scale(sc[, c(5:74)]) #from 26 to 74 +landscape+soils
siteCovs(rbwo.abund) <- sc
landscape1.rbwo <- pcount(~ Jdate + Wind + Sky + Noise + Time
                       ~ Evergreen1km + HighDev1km + Schrubs1km + Ag1km
                          , rbwo.abund, mixture="P", K=120)

NewData.Ag1km  <- data.frame(Evergreen1km=0, Ag1km=seq(min(sc$Ag1km),max(sc$Ag1km), length=100),Schrubs1km=0,HighDev1km=0)
rbwo.est.ag1 <- predict(landscape1.rbwo, type="state",
                              newdata=NewData.Ag1km,appendData=TRUE)

plot(Predicted~ Ag1km, data=rbwo.est.ag1, ylim=c(0,40), type="l", lwd=3,
     xlab="Agricultural cover within 1km of patch", ylab="Est. RBWO Abundance")
##95% confidence intervals
lines(lower~ Ag1km, data=rbwo.est.ag1,  type="l", lwd=3, col="darkgray")
lines(upper~ Ag1km, data=rbwo.est.ag1, type="l", lwd=3, col="darkgray")
```


#RBWO - upstate model but no variables significant

#CAWR - upstate model but no variables significant



# Barplot with error bars if need be

###A function - just copy and paste between the next 6 lines ###
error.bar <- function(x, y, upper, lower=upper, length=0.1,...){
  if(length(x) != length(y) | length(y) !=length(lower) | length(lower) != length(upper))
    stop("vectors must be same length")
  arrows(x,y+upper, x, y-lower, angle=90, code=3, length=length, ...)
}

#### Copy and paste above. Do not edit##

#code below for categorical bar plot data - not yet edited for my data
data.mean<-bhnu.est.matrix$Predicted
data.sd<-bhnu.est.matrix$SE

bar.p <-barplot(data.mean,
                names.arg=c("Semi-urban","Recreation","Protected","Agriculture"),
                ylim = c(0, 10), ylab="Est. BHNU Abundance", xlab="System Type",
                #cex.names = 1.5, cex.axis=1.5, cex.lab=1.5, 
                col="darkblue")

error.bar(bar.p,data.mean,data.sd) #sd
#end 


