---
title: "VEG_figures"
author: "Jesse Wood"
date: "July 19, 2018"
output:
  word_document: 
    fig_width: 5
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##VEG DATA

```{r}
setwd("C:/Users/woodj/Documents/GRAD SCHOOL - CLEMSON/Project-Specific/R work/USDA-songbirds/USDA-songbirds")

vdata <-read.csv("template_pcount_csv.csv")

library("ggplot2")
library("rmarkdown")
library("markdown")
library("knitr")
library("RColorBrewer")
library("lme4")
library("lmtest")
library("unmarked")
library("AICcmodavg")
library("quantreg")

```

#canopy cover varied significantly by treatment
```{r}
ccover.lm<-lm(Ccover~Nburns,data=vdata)
summary(ccover.lm)
confint.lm(ccover.lm)
plot(ccover.lm)


#summary(vdata)
#levels(vdata$Treatment)
#display.brewer.all()
#myColors<-brewer.pal(4,"Greens")
#names(myColors)<-levels(vdata$Treatment)
#colScale<-scale_colour_manual(name="Treatment",values=myColors)
#colScale

#fig+colScale+theme(panel.background = element_rect(fill = "gray",colour = "lightgray",size = 0.5, linetype = "solid"))

#brewer.pal(n=8,name="Greens")
#palette(brewer.pal(n=4,name="Greens"))

fig<-print(ggplot(vdata,aes(x=Nburns,y=Ccover,colour=Treatment))
  +geom_jitter()
  +geom_smooth(method=lm,fullrange=TRUE,color="black",size=0.5)
  +labs(x="Increasing # of burns",y="Canopy cover (more dense)")
  +scale_color_brewer(palette="Greens")+theme(panel.background = element_rect(fill = "tan",colour = "lightgray",size = 0.5, linetype = "solid"))) #or tan1 or lightsalmon1 or navajowhite2
      
#stat_quantile(data=vdata,quantiles=c(0.025,0.975),formula=Ccover~Nburns,method="rq",geom="quantile"))

plot(Ccover~Nburns,data=vdata)
abline(ccover.lm)
```

ignore
```{r}
#1-why are labels and values offset? (NA for 0B)

#vdata$Treatment<-factor(vdata$Treatment, levels=c("OB","1B","2B","3B"),labels=c("OB","1B","2B","3B"))

#confint.lm

#qplot(vdata$Treatment,vdata$Ccover,data=vdata,geom=c("boxplot","jitter"),fill=vdata$Treatment,xlab="Burn Treatments",ylab="Canopy Cover (more to less open)")
```

```{r}
vdata <-read.csv("template_pcount_csv.csv")

#print(ggplot(data=vdata,aes(Treatment,Ccover,colour=Treatment))+geom_boxplot(xlab="Burn Treatments",ylab="Canopy Cover (more to less open)")) #notch=TRUE removed #seems to be same as below right now

quantiles_95<-function(x){
  r<-quantile(x,probs=c(0.025,0.25,0.5,0.75,0.975))
  names(r)<-c("ymin","lower","middle","upper","ymax")
  r
}

ggplot(data=vdata,aes(x=Treatment,y=Ccover,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Canopy cover (more to less open)")

ggplot(data=vdata,aes(x=Nburns,y=Ccover,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Canopy cover (more to less open)")

print(ggplot(data=vdata,aes(Nburns,Ccover))+geom_smooth(xlab="Increasing number of burns",ylab="Canopy cover (more to less open)")+geom_jitter())
```


# relative hardwood to pine in the canopy decr but NOT-SIG by treatment
# YES sig dec by #burns
```{r}
hw2p.lm<-lm(Rel_HW2P_canopy~Nburns,data=vdata)
summary(hw2p.lm)
confint.lm(hw2p.lm)
```

```{r}
fig<-print(ggplot(vdata,aes(x=Nburns,y=Rel_HW2P_canopy,colour=Treatment))
  +geom_jitter()
  +geom_smooth(method=lm,fullrange=TRUE,color="black",size=0.5)
  +labs(x="Increasing # of burns",y="Relative hardwood to pine (in canopy)")
  +scale_color_brewer(palette="Greens")+theme(panel.background = element_rect(fill = "tan",colour = "lightgray",size = 0.5, linetype = "solid"))) #or tan1 or lightsalmon1 or navajowhite2
```


```{r}
#no treatment fig bc not significant

ggplot(data=vdata,aes(x=Nburns,y=Rel_HW2P_canopy,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Relative hardwood to pine (in canopy)")

print(ggplot(data=vdata,aes(Nburns,Rel_HW2P_canopy))+geom_smooth(xlab="Increasing number of burns",ylab="Relative hardwood to pine (in canopy)")+geom_jitter())

```

# litter depth sig decreased for 3B,
# decr sig w Nburns
```{r}
litter.lm<-lm(Ldepth~Nburns,data=vdata)
summary(litter.lm)
confint.lm(litter.lm)
```

```{r}
print(ggplot(vdata,aes(x=Nburns,y=Ldepth,colour=Treatment))
  +geom_jitter()
  +geom_smooth(method=lm,fullrange=TRUE,color="black",size=0.5)
  +labs(x="Increasing # of burns",y="Litter depth (cm)")
  +scale_color_brewer(palette="Greens")+theme(panel.background = element_rect(fill = "tan",colour = "lightgray",size = 0.5, linetype = "solid"))) #or tan1 or lightsalmon1 or navajowhite2
```


```{r}
#qplot(vdata$Nburns,vdata$Ldepth,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Litter depth")

#print(ggplot(data=vdata,aes(Nburns,Ldepth,colour=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Litter depth"))

ggplot(data=vdata,aes(x=Treatment,y=Ldepth,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Litter depth (cm)")

ggplot(data=vdata,aes(x=Nburns,y=Ldepth,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Litter depth (cm)")

print(ggplot(data=vdata,aes(Nburns,Ldepth))+geom_smooth(xlab="Increasing number of burns",ylab="Litter depth (cm)")+geom_jitter())

```

# BA sig decrease for 2B&3B
#   sig decrease w Nburns
```{r}
baT.lm<-lm(BA~Treatment,data=vdata)
summary(baT.lm)
confint.lm(baT.lm)

baB.lm<-lm(BA~Nburns,data=vdata)
summary(baB.lm)
confint.lm(baB.lm)
```

```{r}
fig<-print(ggplot(vdata,aes(x=Nburns,y=BA,colour=Treatment))
  +geom_jitter()
  +geom_smooth(method=lm,fullrange=TRUE,color="black",size=0.5)
  +labs(x="Increasing # of burns",y="Basal Area")
  +scale_color_brewer(palette="Greens")+theme(panel.background = element_rect(fill = "tan",colour = "lightgray",size = 0.5, linetype = "solid"))) #or tan1 or lightsalmon1 or navajowhite2
```


```{r}
#treatment
#qplot(vdata$Treatment,vdata$BA,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Burn treatment",ylab="Basal area")

#print(ggplot(data=vdata,aes(Treatment,BA,colour=Treatment))+geom_boxplot(xlab="Burn treatment",ylab="Basal Area"))

# number of burns
#qplot(vdata$Nburns,vdata$BA,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Basal area")

#print(ggplot(data=vdata,aes(Nburns,BA,colour=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Basal Area"))


#new
ggplot(data=vdata,aes(x=Treatment,y=BA,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Basal area (10-factor)")

ggplot(data=vdata,aes(x=Nburns,y=BA,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Basal area (10-factor)")

print(ggplot(data=vdata,aes(Nburns,BA))+geom_smooth(xlab="Increasing number of burns",ylab="Basal Area")+geom_jitter())

```

# herbaceous cover at low heights sig increase for 2B&3B
#   sig increase w Nburns
```{r}
fgherbT.lm<-lm(FG_herb~Treatment,data=vdata)
summary(fgherbT.lm)
confint.lm(fgherbT.lm)

fgherbB.lm<-lm(FG_herb~Nburns,data=vdata)
summary(fgherbB.lm)
confint.lm(fgherbB.lm)

```

```{r}
fgherbTSB.lm<-lm(FG_herb~TimeSinceB,data=vdata)
summary(fgherbTSB.lm)
confint.lm(fgherbTSB.lm)

plot(FG_herb ~ TimeSinceB,vdata, xlim=c(0,10))
```

```{r}
fig<-print(ggplot(vdata,aes(x=Nburns,y=FG_herb,colour=Treatment))
  +geom_jitter()
  +geom_smooth(method=lm,fullrange=TRUE,color="black",size=0.5)
  +labs(x="Increasing # of burns",y="Herbaceous cover in understory")
  +scale_color_brewer(palette="Greens")+theme(panel.background = element_rect(fill = "tan",colour = "lightgray",size = 0.5, linetype = "solid"))) #or tan1 or lightsalmon1 or navajowhite2
```


```{r}


#treatment
#qplot(vdata$Treatment,vdata$FG_herb,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Burn treatment",ylab="Herbaceous cover in understory")

#colour=c("lightgrey","palegreen","green1","darkgreen")

#print(ggplot(data=vdata,aes(Treatment,FG_herb,fill=Treatment))+geom_boxplot(xlab="Burn treatment",ylab="Herbaceous cover in understory")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter())

# number of burns
#qplot(vdata$Nburns,vdata$FG_herb,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Herbaceous cover in understory")

#print(ggplot(data=vdata,aes(Nburns,FG_herb,fill=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Herbaceous cover in understory")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter())


ggplot(data=vdata,aes(x=Treatment,y=FG_herb,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Herbaceous cover in understory")

ggplot(data=vdata,aes(x=Nburns,y=FG_herb,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Herbaceous cover in understory")

print(ggplot(data=vdata,aes(Nburns,FG_herb))+geom_smooth(xlab="Increasing number of burns",ylab="Herbaceous cover in understory")+geom_jitter())
```

# HW sapling density at low-med heights sig increase for 2B&3B
#   sig increase w Nburns
```{r}
hw1050T.lm<-lm(HW_dens_1050~Treatment,data=vdata)
summary(hw1050T.lm)
confint.lm(hw1050T.lm)

hw1050B.lm<-lm(HW_dens_1050~Nburns,data=vdata)
summary(hw1050B.lm)
confint.lm(hw1050B.lm)
```

```{r}
fig<-print(ggplot(vdata,aes(x=Nburns,y=HW_dens_1050,colour=Treatment))
  +geom_jitter()
  +geom_smooth(method=lm,fullrange=TRUE,color="black",size=0.5)
  +labs(x="Increasing # of burns",y="Hardwood seedlings in understory")
  +scale_color_brewer(palette="Greens")+theme(panel.background = element_rect(fill = "tan",colour = "lightgray",size = 0.5, linetype = "solid"))) #or tan1 or lightsalmon1 or navajowhite2
```

```{r}
hw1050TSB.lm<-lm(HW_dens_1050~TimeSinceB,data=vdata)
summary(hw1050TSB.lm)
confint.lm(hw1050TSB.lm)

plot(HW_dens_1050 ~ TimeSinceB,vdata, xlim=c(0,10))
```

```{r}
#treatment
#qplot(vdata$Treatment,vdata$HW_dens_1050,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Burn treatment",ylab="Hardwood seedling density in understory")

#print(ggplot(data=vdata,aes(Treatment,HW_dens_1050,colour=Treatment))+geom_boxplot(xlab="Burn treatment",ylab="Hardwood seedling density in understory"))

# number of burns
#qplot(vdata$Nburns,vdata$HW_dens_1050,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Hardwood seedling density in understory")

#print(ggplot(data=vdata,aes(Nburns,HW_dens_1050,colour=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Hardwood seedling density in understory"))

ggplot(data=vdata,aes(x=Treatment,y=HW_dens_1050,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Hardwood seedling density in understory")

ggplot(data=vdata,aes(x=Nburns,y=HW_dens_1050,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Hardwood seedling density in understory")

print(ggplot(data=vdata,aes(Nburns,HW_dens_1050))+geom_smooth(xlab="Increasing number of burns",ylab="Hardwood seedling density in understory")+geom_jitter())


```

#FG_shrub non-sig
```{r}
fgshrubB.lm<-lm(FG_shrub~Nburns,data=vdata)
summary(fgshrubB.lm)
confint.lm(fgshrubB.lm)
```

#NHW_saplings sig!
```{r}
HWsaplingsB.lm<-lm(NHW_saplings~Nburns,data=vdata)
summary(HWsaplingsB.lm)
confint.lm(HWsaplingsB.lm)
```

```{r}
fig<-print(ggplot(vdata,aes(x=Nburns,y=NHW_saplings,colour=Treatment))
  +geom_jitter()
  +geom_smooth(method=lm,fullrange=TRUE,color="black",size=0.5)
  +labs(x="Increasing # of burns",y="Hardwood saplings (midstory)")
  +scale_color_brewer(palette="Greens")+theme(panel.background = element_rect(fill = "tan",colour = "lightgray",size = 0.5, linetype = "solid"))) #or tan1 or lightsalmon1 or navajowhite2
```

# snags non-sig
```{r}
snagsB.lm<-lm(Nsnags~Nburns,data=vdata)
summary(snagsB.lm)
confint.lm(snagsB.lm)
```

# snags non-sig
```{r}
snagsB.lm<-lm(Nsnags~Nburns,data=vdata)
summary(snagsB.lm)
confint.lm(snagsB.lm)
```


#thins and age were correlated
```{r}
fig<-plot(Nthins ~ Age, vdata)
abline(lm(vdata$Nthins ~ vdata$Age),col="blue")
cor.test(vdata$Nthins,vdata$Age) #r=0.671
```

```{r}
cor.test(vdata$Nthins,vdata$Nburns) #r=0.506
```

```{r}
thins.lm<-lm(Ccover~Nthins,data=vdata)
summary(thins.lm)
confint.lm(thins.lm)
#plot(thins.lm)
```
```{r}
BAthins.lm<-lm(BA~Nthins,data=vdata)
summary(BAthins.lm)
confint.lm(BAthins.lm)
#plot(thins.lm)
```

```{r}
Hhw1050.lm<-lm(HW_dens_1050~Herbicide,data=vdata)
summary(Hhw1050.lm)
confint.lm(Hhw1050.lm)
```

```{r}
Hhws.lm<-lm(NHW_saplings~Herbicide,data=vdata)
summary(Hhws.lm)
confint.lm(Hhws.lm)
boxplot(NHW_saplings ~ Herbicide, vdata)

quantiles_95<-function(x){
  r<-quantile(x,probs=c(0.025,0.25,0.5,0.75,0.975))
  names(r)<-c("ymin","lower","middle","upper","ymax")
  r
}

ggplot(data=richness,aes(x=Herbicide,y=NHW_saplings,color=Herbicide))+guides(fill=T)+stat_summary(fun.data=quantiles_95,geom="boxplot", color=c("gray40","orange"), labels=c("0","1"))+labs(x="Herbicide",y="# Hardwood Saplings")+coord_cartesian(ylim=c(0, 50)) +
  geom_text(x=0, y=44, label="a",color=1) + geom_text(x=1, y=34, label="b",color=1) +
  theme(panel.border = element_blank(),panel.grid.major =element_blank(), panel.grid.minor = element_blank())


scale_y_continuous(breaks=seq(18, 36, 2))

ggplot(data=richness,aes(x=YearCat,y=SR,fill=YearCat))+scale_fill_manual(name="Survey Year", values=c("red", "blue"),labels=c("2017","2018"))+stat_summary(fun.data=quantiles_95,geom="boxplot")+labs(x="Survey year",y="Species Richness")+coord_cartesian(ylim=c(18, 36)) + scale_y_continuous(breaks=seq(20, 35, 5))+ geom_text(x=1, y=35, label="a",color=1)+ geom_text(x=2, y=31, label="b",color=1) + theme(panel.border = element_blank(),panel.grid.major = element_blank(),panel.grid.minor = element_blank(), legend.position="right", legend.box = "vertical",legend.background=element_blank(),legend.key = element_blank(), )


```



```{r}
Hhwshrub.lm<-lm(HW_shrub~Herbicide,data=vdata)
summary(Hhwshrub.lm)
confint.lm(Hhwshrub.lm)
boxplot(HW_shrub ~ Herbicide, vdata)

ggplot(data=richness,aes(x=Herbicide,y=HW_shrub,color=Herbicide))+guides(fill=T)+
  stat_summary(fun.data=quantiles_95,geom="boxplot", color=c("gray40","orange"), labels=c("0","1"))+labs(x="Herbicide",y="Hardwood cover (midstory)")+coord_cartesian(ylim=c(0, 65)) +
  geom_text(x=0, y=60, label="a",color=1) + geom_text(x=1, y=43, label="b",color=1) +
  theme(panel.border = element_blank(),panel.grid.major =element_blank(), panel.grid.minor = element_blank())
```

#nonsig with FG_shrub or FG_herb

```{r}
Hhw2p.lm<-lm(Rel_HW2P_canopy~Herbicide,data=vdata)
summary(Hhw2p.lm)
confint.lm(Hhw2p.lm)
boxplot(Rel_HW2P_canopy ~ Herbicide, vdata)

ggplot(data=richness,aes(x=Herbicide,y=Rel_HW2P_canopy,color=Herbicide))+guides(fill=T)+
  stat_summary(fun.data=quantiles_95,geom="boxplot", color=c("gray40","orange"), labels=c("0","1"))+labs(x="Herbicide",y="Hardwood to pine ratio")+coord_cartesian(ylim=c(0, 1)) +
  geom_text(x=0, y=.8, label="a",color=1) + geom_text(x=1, y=.6, label="b",color=1) +
  theme(panel.border = element_blank(),panel.grid.major =element_blank(),
        panel.grid.minor = element_blank())
```


NOT using anymore
#boxes/volume
```{r}
plot(Treatment ~ Ccover, data=vdata)
plot(Treatment ~ Rel_HW2P_canopy, data=vdata)
plot(Treatment ~ Ldepth, data=vdata)
plot(Treatment ~ BA, data=vdata)
plot(Treatment ~ FG_herb, data=vdata)
plot(Treatment ~ HW_dens_1050, data=vdata)
```

#simple boxes
```{r}
plot(Ccover ~ Treatment, data=vdata)
plot(Rel_HW2P_canopy ~ Treatment, data=vdata)
plot(Ldepth ~ Treatment, data=vdata)
plot(BA ~ Treatment, data=vdata)
plot(FG_herb ~ Treatment, data=vdata)
plot(HW_dens_1050 ~ Treatment, data=vdata)
```

#scatter, ordered
```{r}
library("ggplot2")
print(ggplot(vdata, aes(Treatment, Ccover)) + geom_point())
print(ggplot(vdata, aes(Treatment, Rel_HW2P_canopy)) + geom_point())
print(ggplot(vdata, aes(Treatment, Ldepth)) + geom_point())
print(ggplot(vdata, aes(Treatment, BA)) + geom_point())
print(ggplot(vdata, aes(Treatment, FG_herb)) + geom_point())
print(ggplot(vdata, aes(Treatment, HW_dens_1050)) + geom_point())
```

#jittered and color-coded treatment (nburns axis) instead
```{r}
print(ggplot(vdata, aes(Nburns, Ccover, colour=Treatment)) + geom_jitter())
print(ggplot(vdata, aes(Nburns, Rel_HW2P_canopy, colour=Treatment)) + geom_jitter())
print(ggplot(vdata, aes(Nburns, Ldepth, colour=Treatment)) + geom_jitter())
print(ggplot(vdata, aes(Nburns, BA, colour=Treatment)) + geom_jitter())
print(ggplot(vdata, aes(Nburns, FG_herb, colour=Treatment)) + geom_jitter())
print(ggplot(vdata, aes(Nburns, HW_dens_1050, colour=Treatment)) + geom_jitter())
```

```{r}
fig<-plot(Nthins ~ Age, vdata)
abline(lm(vdata$Nthins ~ vdata$Age),col="blue")
```
