---
title: "VEG_figures"
author: "Jesse Wood"
date: "July 19, 2018"
output: html_document
output: word_doc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##VEG DATA

```{r}
setwd("C:/Users/woodj/Documents/GRAD SCHOOL - CLEMSON/Project-Specific/R work/USDA-songbirds/USDA-songbirds")

vdata <-read.csv("template_pcount_csv.csv")
library("ggplot2","rmarkdown","markdown","knitr","RColorBrewer","lme4","lmtest","unmarked","AICcmodavg")
```

#canopy cover varied significantly by treatment
```{r}
ccover.lm<-lm(Ccover~Treatment,data=vdata)
summary(ccover.lm)
confint.lm(ccover.lm)
```

ignore
```{r}
1-why are labels and values offset? (NA for 0B)

#vdata$Treatment<-factor(vdata$Treatment, levels=c("OB","1B","2B","3B"),labels=c("OB","1B","2B","3B"))

#confint.lm

#qplot(vdata$Treatment,vdata$Ccover,data=vdata,geom=c("boxplot","jitter"),fill=vdata$Treatment,xlab="Burn Treatments",ylab="Canopy Cover (more to less open)")
```

```{r}
vdata <-read.csv("template_pcount_csv.csv")

#print(ggplot(data=vdata,aes(Treatment,Ccover,colour=Treatment))+geom_boxplot(xlab="Burn Treatments",ylab="Canopy Cover (more to less open)")) #notch=TRUE removed #seems to be same as below right now

quantiles_95<-function(x){
  r<-quantile(x,probs=c(0.025,0.25,0.5,0.75,0.975))
  names(r)<-c("ymin","lower","middle","upper","ymax")
  r
}

ggplot(data=vdata,aes(x=Treatment,y=Ccover,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Canopy cover (more to less open)")

ggplot(data=vdata,aes(x=Nburns,y=Ccover,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Canopy cover (more to less open)")

print(ggplot(data=vdata,aes(Nburns,Ccover))+geom_smooth(xlab="Increasing number of burns",ylab="Canopy cover (more to less open)")+geom_jitter())
```


# relative hardwood to pine in the canopy decr but NOT-SIG by treatment
# YES sig dec by #burns
```{r}
hw2p.lm<-lm(Rel_HW2P_canopy~Nburns,data=vdata)
summary(hw2p.lm)
confint.lm(hw2p.lm)
```

```{r}
#no treatment fig bc not significant

ggplot(data=vdata,aes(x=Nburns,y=Rel_HW2P_canopy,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Relative hardwood to pine (in canopy)")

print(ggplot(data=vdata,aes(Nburns,Rel_HW2P_canopy))+geom_smooth(xlab="Increasing number of burns",ylab="Relative hardwood to pine (in canopy)")+geom_jitter())

```

# litter depth sig decreased for 3B,
# decr sig w Nburns
```{r}
litter.lm<-lm(Ldepth~Nburns,data=vdata)
summary(litter.lm)
confint.lm(litter.lm)
```

```{r}
#qplot(vdata$Nburns,vdata$Ldepth,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Litter depth")

#print(ggplot(data=vdata,aes(Nburns,Ldepth,colour=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Litter depth"))

ggplot(data=vdata,aes(x=Treatment,y=Ldepth,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Litter depth (cm)")

ggplot(data=vdata,aes(x=Nburns,y=Ldepth,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Litter depth (cm)")

print(ggplot(data=vdata,aes(Nburns,Ldepth))+geom_smooth(xlab="Increasing number of burns",ylab="Litter depth (cm)")+geom_jitter())

```

# BA sig decrease for 2B&3B
#   sig decrease w Nburns
```{r}
baT.lm<-lm(BA~Treatment,data=vdata)
summary(baT.lm)
confint.lm(baT.lm)

baB.lm<-lm(BA~Nburns,data=vdata)
summary(baB.lm)
confint.lm(baB.lm)
```

```{r}
#treatment
#qplot(vdata$Treatment,vdata$BA,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Burn treatment",ylab="Basal area")

#print(ggplot(data=vdata,aes(Treatment,BA,colour=Treatment))+geom_boxplot(xlab="Burn treatment",ylab="Basal Area"))

# number of burns
#qplot(vdata$Nburns,vdata$BA,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Basal area")

#print(ggplot(data=vdata,aes(Nburns,BA,colour=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Basal Area"))


#new
ggplot(data=vdata,aes(x=Treatment,y=BA,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Basal area (10-factor)")

ggplot(data=vdata,aes(x=Nburns,y=BA,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Basal area (10-factor)")

print(ggplot(data=vdata,aes(Nburns,BA))+geom_smooth(xlab="Increasing number of burns",ylab="Basal Area")+geom_jitter())

```

# herbaceous cover at low heights sig increase for 2B&3B
#   sig increase w Nburns
```{r}
fgherbT.lm<-lm(FG_herb~Treatment,data=vdata)
summary(fgherbT.lm)
confint.lm(fgherbT.lm)

fgherbB.lm<-lm(FG_herb~Nburns,data=vdata)
summary(fgherbB.lm)
confint.lm(fgherbB.lm)
```

```{r}


#treatment
#qplot(vdata$Treatment,vdata$FG_herb,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Burn treatment",ylab="Herbaceous cover in understory")

#colour=c("lightgrey","palegreen","green1","darkgreen")

#print(ggplot(data=vdata,aes(Treatment,FG_herb,fill=Treatment))+geom_boxplot(xlab="Burn treatment",ylab="Herbaceous cover in understory")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter())

# number of burns
#qplot(vdata$Nburns,vdata$FG_herb,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Herbaceous cover in understory")

#print(ggplot(data=vdata,aes(Nburns,FG_herb,fill=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Herbaceous cover in understory")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter())


ggplot(data=vdata,aes(x=Treatment,y=FG_herb,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Herbaceous cover in understory")

ggplot(data=vdata,aes(x=Nburns,y=FG_herb,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Herbaceous cover in understory")

print(ggplot(data=vdata,aes(Nburns,FG_herb))+geom_smooth(xlab="Increasing number of burns",ylab="Herbaceous cover in understory")+geom_jitter())
```

# HW sapling density at low-med heights sig increase for 2B&3B
#   sig increase w Nburns
```{r}
hw1050T.lm<-lm(HW_dens_1050~Treatment,data=vdata)
summary(hw1050T.lm)
confint.lm(hw1050T.lm)

hw1050B.lm<-lm(HW_dens_1050~Nburns,data=vdata)
summary(hw1050B.lm)
confint.lm(hw1050B.lm)
```

```{r}
#treatment
#qplot(vdata$Treatment,vdata$HW_dens_1050,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Burn treatment",ylab="Hardwood seedling density in understory")

#print(ggplot(data=vdata,aes(Treatment,HW_dens_1050,colour=Treatment))+geom_boxplot(xlab="Burn treatment",ylab="Hardwood seedling density in understory"))

# number of burns
#qplot(vdata$Nburns,vdata$HW_dens_1050,data=vdata,geom=c("boxplot","jitter"),color=vdata$Treatment,xlab="Increasing number of burns",ylab="Hardwood seedling density in understory")

#print(ggplot(data=vdata,aes(Nburns,HW_dens_1050,colour=Treatment))+geom_boxplot(xlab="Increasing number of burns",ylab="Hardwood seedling density in understory"))

ggplot(data=vdata,aes(x=Treatment,y=HW_dens_1050,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Burn treatment",y="Hardwood seedling density in understory")

ggplot(data=vdata,aes(x=Nburns,y=HW_dens_1050,fill=Treatment))+guides(fill=F)+stat_summary(fun.data=quantiles_95,geom="boxplot")+scale_fill_brewer(type="seq",direction=1,palette="Greens")+geom_jitter()+labs(x="Increasing # of burns",y="Hardwood seedling density in understory")

print(ggplot(data=vdata,aes(Nburns,HW_dens_1050))+geom_smooth(xlab="Increasing number of burns",ylab="Hardwood seedling density in understory")+geom_jitter())


```

```{r}
fig<-plot(Nthins ~ Age, vdata)
abline(lm(vdata$Nthins ~ vdata$Age),col="blue")
```